### 2020-11-27 latex/acrobot.tex
Completed the edit of oscillation proof

In the spirit of perturbation analysis, we will expand our time-scaled system
\(\hat{\mu}(\alpha,\mu_0,I)\).
We begin by setting \(I = 0\) to find the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\hat{\mu}_0,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
   .
\end{cases}
\]
Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom})
imply that \(g(\alpha,\mu,0) = 0\),
so the solution to this nominal system is 
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

By ~\eqref{eqn:khalil-perturbation-firstorder}, we know the function
\(\hat{\mu}_1(\alpha,\mu_0)\) is the solution to the linear time-varying scalar
system
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = 
        \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)\hat{\mu}_1 + \pdiff{g}{I}(\alpha, \mu_0, 0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
These dynamics are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0) &= 0
    , \\
    \pdiff{g}{I}(\alpha, \mu_0, 0) &= K b(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    b(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
    .
\]
Therefore, the first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\) is
\begin{equation}\label{eqn:acrobot-muhat-expanded}
    \hat{\mu}(\alpha,\mu_0,I)
    = \mu_0 + I K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma +
    R(\alpha,\mu_0,I)
    .
\end{equation}

@@

%Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
%After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
%negative \(q_u\)-axis with pseudo-radius 
%\(u = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
%Hence, the change in pseudo-radius is 
%\(u - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
%If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
%it will land on the positive \(q_u\)-axis
%with pseudo-radius
%\(\tilde{u} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
%
%Notice that \(b(\alpha,\mu_0)\) is even and 
%\(\pi\)-periodic in \(\alpha\), so that
%\begin{equation}\label{eqn:bhat-integral-properties}
%    \int \limits_{-\pi}^\pi b(\sigma,\mu_0) d\sigma
%    = \int \limits_0 ^ {2 \pi} b(\sigma,\mu_0) d\sigma
%    = 2 \int \limits_0^\pi b(\sigma,\mu_0) d\sigma
%    .
%\end{equation}
%This implies that 
%\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
%which means the change in pseudo-radius after a full oscillation is 
%\(\tilde{u} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
%If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
%then both \(u - \mu_0 > 0\) and \(\tilde{u} - \mu_0 > 0\).
%In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
%the behaviour at each half oscillation.

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system, we can define this Poincar\'{e} map as the change
in \(\hat{\mu}\) at each of these intersections:
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
From our perturbation analysis, the section expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi b(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi b(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters, while \(b(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-10}, \pi - 10^{-3}]\)
in Figure \ref{fig:acrobot-Q}. 
We see that it is strictly positive and monotonically increasing, with
an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) result in an infinite integration error,
so we believe that \(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

@@

The Poincar\'{e} section \(P_\mathcal{O}\) allows us understand the evolution of
\(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
\begin{equation}\label{eqn:muhat-discrete}
    \begin{cases}
        \mu(n+1) = P_\mathcal{O}(\mu(n)) 
        = \mu(n) + I K Q(\mu(n)) + R(\pi,\mu(n),I)
        , \\
        \mu(0) = \mu_0
        .
    \end{cases}
\end{equation}
Tolerating the abuse of notation, \(\mu(n)\) represents the distance along the
\(q_u\)-axis when an orbit of the constrained dynamics intersects the
\(q_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (\mu_0,0)\).

Show energy gain on \(\mathcal{O}\) is equivalent to 
finding some \(I^\star \in ]0, I_1]\) where, for all 
\(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
for all \(n \geq N\), \(\mu(n) \notin ]0,\pi-\delta]\).
We will show \(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),
which implies this characterization of energy gain holds true.

Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu = 0\) yields
\[
    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
    \pdiff{R}{\mu}(\pi,0,I)
    .
\]
If there is some value of \(I \in ]0,I_1]\) for which this term is greater than
\(1\), then \(0\) is a repeller of the discrete time system.
To show such a value of \(I\) exists, we first compute 
\[
    Q^\prime(0) = \int \limits_0^\pi \pdiff{b}{\mu}(\sigma,0) d\sigma
    .
\]
MATLAB reveals that 
\[
    \lim \limits_{\mu_0 \to 0^+}
    \pdiff{b}{\mu}(\alpha,\mu_0) 
    = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
    ,
\]
which means 
\[
    Q^\prime(0)  = \frac{\pi}{2\sqrt{2}} > 0
    .
\]
Since the remainder term \(R(\alpha,\mu_0,I)\) is \(O(I^2)\), 
its partial derivative \(\pdiff{R}{\mu}(\pi,\mu_0,I)\) is also \(O(I^2)\).
Hence, it can be written in the form
\(I^2 \tilde{R}^\prime(\mu_0,I)\) where \(\tilde{R}^\prime(\mu_0,I)\) is smooth and
zero at \(I = 0\).
Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly;
thus, there exists \(I_2 \in ]0,I_1]\) where 
\(I_2 K Q^\prime(0) + I_2^2 \tilde{R}^\prime(0,I) > 0\).
Hence,
\[
    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}^\prime(0,I) 
    > 1
    ,
\] 
for all \(I \in ]0, I_2]\), .

We have shown \(0\) is a repeller of the discrete time system, which means
there exists some (unknown) \(\epsilon > 0\) where the interval
\(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
All solutions starting in this interval will flow towards the value
\(\mu = \epsilon\).
In other words, we have found a value \(I_2 > 0\) small enough
that \(\mu(n)\) is strictly increasing on \([0,\epsilon[\).

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in ]0, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
There clearly exists a small enough \(I^\star \in ]0,I_3]\) making
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
\[
    P_\mathcal{O}(\mu_0) > \mu_0
    ,
\]
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
there exists \(I > 0)\) small enough that the constraint 
~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).

Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
monotonically decreasing in \(\mu_0\).
By the same arguments presented in this section, the constraint 
must be dissipating energy on \(\mathcal{O}\).

>>>>

In the spirit of perturbation analysis, we will expand our time-scaled system
\(\hat{\mu}(\alpha,\mu_0,I)\).
We begin by setting \(I = 0\) to find the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\hat{\mu}_0,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
   .
\end{cases}
\]
Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom})
reveal that \(g(\alpha,\mu,0) = 0\),
so the solution to this nominal system is 
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

By ~\eqref{eqn:khalil-perturbation-firstorder}, we know the function
\(\hat{\mu}_1(\alpha,\mu_0)\) is the solution to the linear time-varying scalar
system
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = 
        \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)\hat{\mu}_1 + \pdiff{g}{I}(\alpha, \mu_0, 0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
These dynamics are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0) &= 0
    , \\
    \pdiff{g}{I}(\alpha, \mu_0, 0) &= K a(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha a(\sigma,\mu_0)d\sigma
    .
\]
Therefore, the first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\) is
\begin{equation}\label{eqn:acrobot-muhat-expanded}
    \hat{\mu}(\alpha,\mu_0,I)
    = \mu_0 + I K \int \limits_0^\alpha a(\sigma,\mu_0)d\sigma +
    R(\alpha,\mu_0,I)
    .
\end{equation}

@@

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system, we can define this Poincar\'{e} map as the change
in \(\hat{\mu}\) when \(\alpha\) increases by \(\pi\):
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
From our perturbation analysis, the Poincar\'{e} map -- also known as a
Poincar\'{e} section --  expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters, while \(a(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-10}, \pi - 10^{-3}]\)
in Figure \ref{fig:acrobot-Q}. 
We see that it is strictly positive and monotonically increasing, with
an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) result in an infinite integral error due
to a division by zero, so we believe that \(Q(\mu_0)\) is in fact positive for
all \(\mu_0\).

@@

The Poincar\'{e} section \(P_\mathcal{O}\) allows us understand the evolution of
\(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
system
\begin{equation}\label{eqn:muhat-discrete}
    \begin{cases}
        \mu(n+1) := P_\mathcal{O}(\mu(n)) 
        = \mu(n) + I K Q(\mu(n)) + R(\pi,\mu(n),I)
        , \\
        \mu(0) = \mu_0
        .
    \end{cases}
\end{equation}
Tolerating the abuse of notation, \(\mu(n)\) represents the distance along the
\(q_u\)-axis when an orbit of the constrained dynamics intersects the
\(q_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (\mu_0,0)\).

Show energy gain on \(\mathcal{O}\) is equivalent to showing \(\mu(n)\)
eventually reaches \(\pi\).
That is, we want to find \(I^\star \in ]0, I_1]\) where, for all 
\(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
for all \(n \geq N\), \(\mu(n) \notin [0,\pi-\delta]\).
A sufficient condition for this characterization is to show 
\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),

Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu = 0\) yields
\[
    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
    \pdiff{R}{\mu}(\pi,0,I)
    .
\]
If there is some value of \(I \in ]0,I_1]\) for which this term is greater than
\(1\), then \(0\) is a repeller of the discrete time system.
To show such a value of \(I\) exists, we first compute 
\[
    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
    .
\]
MATLAB reveals that 
\[
    \lim \limits_{\mu_0 \to 0^+}
    \pdiff{a}{\mu}(\alpha,\mu_0) 
    = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
    ,
\]
which means 
\[
    Q^\prime(0)  = \frac{\pi}{2\sqrt{2}} > 0
    .
\]
Since the remainder term \(R(\alpha,\mu_0,I)\) is \(O(I^2)\), 
its partial derivative \(\pdiff{R}{\mu}(\pi,\mu_0,I)\) is also \(O(I^2)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(\mu_0,I)\) where \(\tilde{R}(\mu_0,I)\) is smooth and
zero at \(I = 0\).
Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly;
thus, there exists \(I_2 \in ]0,I_1]\) where 
\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(0,I) > 0\).
Hence,
\[
    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(0,I) 
    > 1
    ,
\] 
for all \(I \in ]0, I_2]\), .

We have shown \(0\) is a repeller of the discrete time system, which means
there exists some (unknown) \(\epsilon > 0\) where the interval
\(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
What's more, all solutions starting in this interval will flow towards the value
\(\mu = \epsilon\).
In other words, we have found a value \(I_2 > 0\) small enough
that \(\mu(n)\) is strictly increasing on \(]0,\epsilon[\).

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
There exists a small enough \(I^\star \in ]0,I_3]\) making
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
\[
    P_\mathcal{O}(\mu_0) > \mu_0
    ,
\]
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
there exists \(I > 0\) small enough that the constraint 
~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).

Setting \(I < 0\), the Poincar\'{e} section now satisfies
\[
    P_\mathcal{O}(\mu_0) < \mu_0
    ,
\]
thanks to the same arguments presented in this section. 
The constraint is therefore dissipating energy on \(\mathcal{O}\).
